---
title: "Exploratory Data Analysis"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)

b_dat = read_csv("CollegeBballPlayers2009-2021.csv")
head(b_dat, 5)

```



### q1: What statistic out of points, rebounds, assists, blocks, and steals has the highest correlation with being drafted?

```{r}
library(corrplot)

# Creating dataframe of desired variables

Statistics <- select(b_dat, pick, pts, treb, ast, blk, stl)

# Corrplot removing NA's for pick variable

corrplot(cor(Statistics[complete.cases(Statistics[,1:6]),]), type = "upper")


#Based on this plot, none of the statistics in question have a strong correlation with the variable "picks". The statistic with the highest correlation is #steals. It is important to notice however that steals are highly correlated with both points and assists. While the impact this has on the answer to this #question is not really known, there may be other factors that impact why players with a lot of steals tend to get drafted at a higher rate.

```


### Q2: Did the UNC team that had the highest effective field goal percentage have more players drafted than the UNC team in 2017 who won a national championship?


```{r}
players_50_percent = filter(b_dat, eFG < 50 & mp > 0 )

players_50_percent_distinct = players_50_percent %>% distinct(player_name, .keep_all= TRUE)

players_50_percent_distinct
```

Now that I have created an appropriate dataset, I will create a new dataset of players from the "players_50_percent" that have been drafted (pick != 0).

```{r}
players_under_50_drafted = filter(players_50_percent_distinct, pick != 0) 
  

players_under_50_drafted_compact = select(players_under_50_drafted, player_name, team, GP, eFG, pick)


kable(players_under_50_drafted_compact)
```
First I will filter players who have an effective field goal percentage of less than 50%, and then I will remove players who have played zero minutes as they might affect the analysis. I will also remove any duplicate players, since any player who played more than one year in college will have more than one corresponding observation. If a player was drafted, their corresponding pick number shows up in all of their respective observations.


```{r}
players_under_50_percent = filter(b_dat, eFG < 50 & mp > 0 )

players_50_percent_distinct = players_under_50_percent %>% distinct(player_name, .keep_all= TRUE)

players_50_percent_distinct
```
Now that I have created an appropriate dataset, I will create a new dataset of players from the "players_50_percent" that have been drafted (pick != 0).



```{r}
players_under_50_drafted = filter(players_50_percent_distinct, pick != 0) 
  

players_under_50_drafted_compact = select(players_under_50_drafted, player_name, team, GP, eFG, pick)


kable(players_under_50_drafted_compact)

```

There are 276 players who played college basketball from 2009-2021 who have had an effective field goal percentage of less than 50% and have been drafted in the NBA draft. This means that 1.35% (calculated by 276/20,328) of players who have an effective field goal percentage of less than 50% have been drafted between 2009-2021.


Now I will look at the players who had an eFG of over 50% and see how many were drafted.

```{r}
# Players who had eFG > 50%

players_over_50_percent = filter(b_dat, eFG > 50 & mp > 0 )

players_over_50_percent_distinct = players_over_50_percent %>% distinct(player_name, .keep_all= TRUE)

head(players_over_50_percent_distinct)


# Players who had eFG > 50% and were drafted

players_over_50_drafted = filter(players_over_50_percent_distinct, pick != 0) 
  

players_over_50_drafted_compact = select(players_over_50_drafted, player_name, team, GP, eFG, pick)


kable(players_over_50_drafted_compact)

```
There are 538 players who played college basketball from 2009-2021 who have had an effective field goal percentage of greater than 50% and have been drafted in the NBA draft. This means that 4.008% (calculated by 538/13423) of players who have an effective field goal percentage of greater than 50% have been drafted between 2009-2021.


```{r}
# Create Pie Chart
under_50_percentage = (nrow(players_under_50_drafted_compact))

over_50_percentage = (nrow(players_over_50_drafted_compact))

slices <- c(under_50_percentage, over_50_percentage)
lbls <- c("eFG under 50%", "eFG over 50%")
pie(slices, labels = lbls, main="Percent of Players Drafted to the NBA based on eFG")

```


In the 2009 NCAA basketball season, The Big XII leads in average ORTG, the Patriot league leads in average DRTG, the Big 10 leads in average ADRTG, and the Big XII leads in average BPM.

### Q1: Which Conference led in Ortg, Drtg, Adrtg, and bpm in 2009?

```{r}
sheet_2009 = filter(b_dat, year == 2009)
conference_sheet_2009 = group_by(sheet_2009, conf)
conference_stats = conference_sheet_2009 %>%
  summarise(ORTG = mean(Ortg), DRTG = mean(drtg), ADRTG = mean(adrtg), BPM = mean(bpm)) 

kable(head(arrange(conference_stats, desc(ORTG))))
kable(head(arrange(conference_stats, DRTG)))
kable(head(arrange(conference_stats, ADRTG)))
kable(head(arrange(conference_stats, desc(BPM))))
```
In the 2009 NCAA basketball season, The Big XII leads in average ORTG, the Patriot league leads in average DRTG, the Big 10 leads in average ADRTG, and the Big XII leads in average BPM. 

### Q2: Which Clemson player had the worst BPM from 2009-2021?

```{r}

clemson = filter(b_dat, team == "Clemson")
clemson$cumulativeBPM = clemson$bpm * clemson$GP
clemson = clemson %>%
  group_by(player_name) %>%
  summarise(BPM = sum(cumulativeBPM))
head(arrange(clemson, BPM))
```

Josh Smith (2011-2015) is the player on Clemson with the worst box plus minus in the last 15 years.

### Q1: Which conference had the highest effective field goal percentage in 2018?

```{r}
eFG_conf <- b_dat %>% 
  filter(year == 2018) %>%
  group_by(conf) %>%
  summarise(eFG = mean(eFG)) %>%
  arrange(-eFG)

eFG_conf
```

In 2018, the America East Conference had the highest effective field goal percentage with 48.46%.

### Q2: What will the correlation between usage rate and year be, charting four year players?

```{r}

four_year_players <- b_dat %>%
  select(pid, player_name, team, yr) %>%
  group_by(pid) %>%
  count(pid) %>%
  filter(n==4) %>%
  inner_join(b_dat, by=c("pid")) %>%
  select(pid, player_name, team, yr)
#these are players that appear in the data set four times ^

four_year_players <- four_year_players %>%
  group_by(pid, team) %>%
  count(pid, team) %>%
  filter(n==4) %>%
  inner_join(b_dat, by=c("pid", "team")) %>%
  select(pid, player_name, team, yr)
#these are players that played all four years with the same team ^

four_year_players <- four_year_players %>%
  group_by(pid, yr) %>%
  count(pid, yr) %>%
  pivot_wider(names_from = "yr", values_from = "n") %>%
  filter(Fr == 1, So == 1, Jr == 1, Sr == 1) %>%
  select(pid)
#these are players whose four years are traditionally labeled as freshman through senior

four_year_usg <- four_year_players %>%
  inner_join(b_dat, by=c("pid")) %>%
  select(pid, player_name, team, yr, usg) %>%
  pivot_wider(names_from = "yr", values_from = "usg") %>%
  ungroup()

#this is a table of every four year player and their usage stats freshman through senior year

means<- four_year_usg %>%
  select(4:7) %>%
  colMeans(na.rm = TRUE)


library(tibble)
means <- means %>% as.data.frame() %>%
  tibble::rownames_to_column("Yr") %>%
  rename("usg" = 2)
#this code converts the named list into a data frame with the average usg rates

#On average, four-year college players should expect their usage rate to grow from 17.40 freshman year to 19.23 senior year.

#Plotting usg data

x <- factor(means$Yr, levels = c("Fr", "So", "Jr", "Sr"))
y <- means$usg

plot(x, y, xlab="Year in School", ylab="Mean Usage Rate")

#Running significance test for usg data
fr_usg <- four_year_usg$Fr
so_usg <- four_year_usg$So
jr_usg <- four_year_usg$Jr
sr_usg <- four_year_usg$Sr

t.test(sr_usg, fr_usg, alternative = "two.sided", var.equal = FALSE)
t.test(so_usg, fr_usg, alternative = "two.sided", var.equal = FALSE)
t.test(jr_usg, so_usg, alternative = "two.sided", var.equal = FALSE)

#These two-sided difference of means tests show that there is a significant difference between the mean usg rates of freshmen and senior players (test 1). There #is also a significant difference between freshmen and sophomores (test 2) and sophomores and juniors (test 3).

```

On average, four-year college players should expect their usage rate to grow from 17.40 freshman year to 19.23 senior year.

### Q1: Which player scored the most points in their college career out of all players from 2009-2021, and were they drafted in the first round?


```{r}
library(dplyr)
library(tidyverse)

b_dat=mutate(b_dat,totalpts=pts*GP)
b_dat=mutate(b_dat, totalptsrnd=round(totalpts,0))
Boypointstotal= b_dat %>%
  group_by(player_name, team, pick) %>%
  summarise(TOTALPTS =sum(totalptsrnd))
library(knitr)
kable(head(arrange(Boypointstotal, desc(TOTALPTS))))

```

### Q2: Did the UNC team that had the highest effective field goal percentage have more players drafted than the UNC team in 2017 who won a national championship?

```{r}
UNCefG<-filter(b_dat, team=="North Carolina")
UNCefG<-UNCefG[, c(4,8,18,21,32,46)]
UNCefG<-mutate(UNCefG, weight=GP*twoPA+TPA)
UNC2009<-filter(UNCefG, year=="2009")
MEAN09<-weighted.mean(UNC2009$eFG, UNC2009$weight)
UNC2010<-filter(UNCefG, year=="2010")
MEAN10<-weighted.mean(UNC2010$eFG, UNC2010$weight)
UNC2011<-filter(UNCefG, year=="2011")
MEAN11<-weighted.mean(UNC2011$eFG, UNC2011$weight)
UNC2012<-filter(UNCefG, year=="2012")
MEAN12<-weighted.mean(UNC2012$eFG, UNC2012$weight)
UNC2013<-filter(UNCefG, year=="2013")
MEAN13<-weighted.mean(UNC2013$eFG, UNC2013$weight)
UNC2014<-filter(UNCefG, year=="2014")
MEAN14<-weighted.mean(UNC2014$eFG, UNC2014$weight)
UNC2015<-filter(UNCefG, year=="2015")
MEAN15<-weighted.mean(UNC2015$eFG, UNC2015$weight)
UNC2016<-filter(UNCefG, year=="2016")
MEAN16<-weighted.mean(UNC2016$eFG, UNC2016$weight)
UNC2017<-filter(UNCefG, year=="2017")
MEAN17<-weighted.mean(UNC2017$eFG, UNC2017$weight)
UNC2018<-filter(UNCefG, year=="2018")
MEAN18<-weighted.mean(UNC2018$eFG, UNC2018$weight)
UNC2019<-filter(UNCefG, year=="2019")
MEAN19<-weighted.mean(UNC2019$eFG, UNC2019$weight)
UNC2020<-filter(UNCefG, year=="2020")
MEAN20<-weighted.mean(UNC2020$eFG, UNC2020$weight)
table(UNC2016$pick)
table(UNC2017$pick)

MEANS<-c(MEAN09,MEAN10,MEAN11,MEAN12,MEAN13,MEAN14,MEAN15,MEAN16,MEAN17,MEAN18,MEAN19,MEAN20)
plot(MEANS, type="o", col="blue",xlab="2009-2020 eFG")

```


### Q1: From 2009-2021, how have the median points per game for freshmen-Seniors changed?

```{r}

library(ggplot2)

b_dat2 <- b_dat[c("pts", "yr", "year")]
#head(b_dat2)

b_dat3 <- filter(b_dat2, yr == "Fr")
#head(b_dat3)

bdat6<- group_by(b_dat3, year)
head(bdat6)
 
Frmean <- bdat6%>%
  summarise(Median_Years = median(pts, na.rm = TRUE))

head(Frmean, 13)

ggplot(data=Frmean, aes(x= year, y= Median_Years)) +
  geom_bar(stat = "identity", fill = "steelblue")+
   geom_text(aes(label = Median_Years), vjust = 2, size = 3)+
  labs(title="FRESHMAN",
        x ="YEARS", y = "MEDIAN PTS")

# Choosing to directly analyze freshman median points in order to see their direct impact on the game in regards to points and whether or not 
# over a span of 13 yrs if we have seen a greater presence of freshman in college basketball.


#Sophomores
b2 <- b_dat[c("pts", "yr", "year")]
#head(b_dat2)

b3 <- filter(b2, yr == "So")
#head(b_dat3)

b4<- group_by(b3, year)

 
Somean <- b4%>%
  summarise(Median_Years = median(pts, na.rm = TRUE))


ggplot(data=Somean, aes(x= year, y= Median_Years)) +
  geom_bar(stat = "identity", fill = "steelblue")+
   geom_text(aes(label = Median_Years), vjust = 2, size = 3)+
  labs(title="SOPHOMORES",
        x ="YEARS", y = "MEDIAN PTS")

# Juniors
b5 <- b_dat[c("pts", "yr", "year")]
#head(b_dat2)

b6 <- filter(b5, yr == "Jr")
#head(b_dat3)

b7<- group_by(b6, year)
head(bdat6)
 
Jrmean <- b7%>%
  summarise(Median_Years = median(pts, na.rm = TRUE))

head(Jrmean, 13)

ggplot(data=Jrmean, aes(x= year, y= Median_Years)) +
  geom_bar(stat = "identity", fill = "steelblue")+
   geom_text(aes(label = Median_Years), vjust = 2, size = 3)+
  labs(title="JUNIORS",
        x ="YEARS", y = "MEDIAN PTS")



#Seniors
b8 <- b_dat[c("pts", "yr", "year")]
#head(b_dat2)

b9 <- filter(b8, yr == "Sr")
#head(b_dat3)

b9<- group_by(b9, year)
head(bdat6)
 
Srmean <- b9%>%
  summarise(Median_Years = median(pts, na.rm = TRUE))

head(Srmean, 13)

ggplot(data=Srmean, aes(x= year, y= Median_Years)) +
  geom_bar(stat = "identity", fill = "steelblue")+
   geom_text(aes(label = Median_Years), vjust = 2, size = 3)+
  labs(title="SENIORS",
        x ="YEARS", y = "MEDIAN PTS")

# from this data we are able to see that as the player progresses from year to year they their impact on this game in regards to points increases. 

```

### Q2: How did that team's effective field goal percentage compare to the 2017 UNC team's effective field goal percentage?

```{r}
#2009 team efg, highest efg
b_dat10 <- b_dat[c("eFG", "team", "year")]
#head(b_dat10)

b_dat11 <- filter(b_dat10, team == "North Carolina")
#head(b_dat11)

b_dat12 <- filter(b_dat11, year== "2009")
#head(b_dat12)
mean09<- mean(b_dat12$eFG)
mean9<- b_dat12$eFG
head(mean9)



#2017 team efg, championship year
b_dat13 <- b_dat[c("eFG", "team", "year")]
#head(b_dat10)

b_dat14 <- filter(b_dat13, team == "North Carolina")
#head(b_dat14)

b_dat15 <- filter(b_dat14, year== "2017")
#head(b_dat15)

mean17<- mean(b_dat15$eFG)
mean7<- b_dat15$eFG


MEANS<-c(mean7, mean9)
plot(MEANS, type="o", col="red")



#Given the two sample t-test we are able to take note that there is no statistical significance between the two years since our p-val was great than .05. 
t.test(mean9 , mean7, alternative = "two.sided", var.equal = FALSE)



```

# Follow-up Questions

### New Questions Based Off Initial Investigation

-   Q1: Which college had the most players over the 13 year span? Which college had the least number of players over the 13 year span & what was the distribution?


-   Q2: By examining statistics such as three point shots attempted, offensive rebounds, assists, etc., can we make an inference about the play style of certain conferences, i.e. does the ACC tend to play at a fast pace, or does the BIG10 rely heavily on the three point shot, and is there a correlation between this determined play style and the number of players drafted to the NBA who play in these systems?


### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE


SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.

```{r}
#Q1:
  
players_per_team = group_by(b_dat, team, player_name)
players = count(players_per_team, team)
players = select(players, c(1,2))
number_per_team = group_by(players, team)
number_per_team = summarise(number_per_team, count = n())
number_per_team

kable(head(arrange(number_per_team, desc(count))))
kable(head(arrange(number_per_team, count)))
ggplot(data = number_per_team) +
  geom_histogram(mapping = aes(x = count), binwidth = 2)
```
Mississippi Valley State is the school with the most players over the last 13 years, while UC San Diego is the school with the least. The data seems to be left-skewed, but this due to the fact that some schools don't have data for some of the years, since they were not Division I NCAA schools for basketball. Removing these outliers, the distribution of the data seems to be fairly even without much skewness. . 



3.  By examining statistics such as three point shots attempted, offensive rebounds, assists, etc., can we make an inference about the play style of certain conferences, i.e. does the ACC tend to play at a fast pace, or does the BIG10 rely heavily on the three point shot, and is there a correlation between this determined play style and the number of players drafted to the NBA who play in these systems?


```{r}
#Q2:
# Data frame of all players drafted

drafted = filter(b_dat, is.na(pick) == FALSE)

# Remove repeat names

drafted_distinct = drafted %>% distinct(player_name, .keep_all= TRUE)


# Count picks by conference

picks_by_conference = count(drafted_distinct, conf)


# Seven conferences with most total draft picks

ggplot(picks_by_conference) +
  geom_col(mapping = aes(x = n, y = conf, fill = conf))
  
```

The six conferences with the highest number of draft picks are the ACC, BIG10, Big 12, Big East, PAC 12, and SEC. For each conference, we will compute the conference wide totals for the following statistics: three point shots attempted, three point percentage, offensive rebounds, free throws attempted, free throws made, blocks per game, and steals per game. We will then compare these totals with those from the other conferences to try and infer certain play styles for these conferences. For percentage based statistics I will compute the median.

```{r}
# ACC Analysis
conference_breakdown = select(b_dat, conf, team, player_name, twoPM, twoPA, TPA, TPM, TP_per, blk_per, stl_per, FTA, FTM, FT_per, oreb, dreb, ast, stl, blk, pts)

ACC_breakdown = filter(conference_breakdown, conf == "ACC")

ACC_breakdown

# Total Shots Attempted

ACC_TSA = sum(ACC_breakdown$twoPA + ACC_breakdown$TPA)

# Three point shots attempted

ACC_tpa_total = sum(ACC_breakdown$TPA)

ACC_tpa_total

# Reliance on three point shot 

ACC_reliance = ACC_tpa_total / ACC_TSA

ACC_reliance * 100

# Three point percentage

ACC_TPper_total = median(ACC_breakdown$TP_per)

ACC_TPper_total * 100

# Offensive rebounds

ACC_oreb_total = sum(ACC_breakdown$oreb)

ACC_oreb_total

# Free throws attempted

ACC_FTA_total = sum(ACC_breakdown$FTA)

ACC_FTA_total

# Free throws made

ACC_FTper_total = median(ACC_breakdown$FT_per)

ACC_FTper_total * 100

# Blocks per game

ACC_BLKper_total = median(ACC_breakdown$blk_per)

ACC_BLKper_total 

# Steals per game

ACC_STLper_total = median(ACC_breakdown$stl_per)

ACC_STLper_total 

```
```{r}
# BIG10 Analysis

# Create desired data frame

B10_breakdown = filter(conference_breakdown, conf == "B10")


# Total Shots Attempted

B10_TSA = sum(B10_breakdown$twoPA + B10_breakdown$TPA)

# Three point shots attempted

B10_tpa_total = sum(B10_breakdown$TPA)

B10_tpa_total

# Reliance on three point shot 

B10_reliance = B10_tpa_total / B10_TSA

B10_reliance * 100

# Three point percentage

B10_TPper_total = median(B10_breakdown$TP_per)

B10_TPper_total * 100

# Offensive rebounds

B10_oreb_total = sum(B10_breakdown$oreb, na.rm = TRUE)

B10_oreb_total

# Free throws attempted

B10_FTA_total = sum(B10_breakdown$FTA)

B10_FTA_total

# Free throws made

B10_FTper_total = median(B10_breakdown$FT_per)

B10_FTper_total * 100

# Blocks per game

B10_BLKper_total = median(B10_breakdown$blk_per)

B10_BLKper_total 

# Steals per game

B10_STLper_total = median(B10_breakdown$stl_per)

B10_STLper_total 

```



```{r}
# Big 12 Analysis

# Create desired data frame

B12_breakdown = filter(conference_breakdown, conf == "B12")

B12_breakdown


# Total Shots Attempted

B12_TSA = sum(B12_breakdown$twoPA + B12_breakdown$TPA)

# Three point shots attempted

B12_tpa_total = sum(B12_breakdown$TPA)

B12_tpa_total

# Reliance on three point shot 

B12_reliance = B12_tpa_total / B12_TSA

B12_reliance * 100

# Three point percentage

B12_TPper_total = median(B12_breakdown$TP_per)

B12_TPper_total * 100

# Offensive rebounds

B12_oreb_total = sum(B12_breakdown$oreb, na.rm = TRUE)

B12_oreb_total

# Free throws attempted

B12_FTA_total = sum(B12_breakdown$FTA)

B12_FTA_total

# Free throws made

B12_FTper_total = median(B12_breakdown$FT_per)

B12_FTper_total * 100

# Blocks per game

B12_BLKper_total = median(B12_breakdown$blk_per)

B12_BLKper_total 

# Steals per game

B12_STLper_total = median(B12_breakdown$stl_per)

B12_STLper_total 
```
```{r}
# Big East Analysis

# Create desired data frame

BE_breakdown = filter(conference_breakdown, conf == "BE")

BE_breakdown


# Total Shots Attempted

BE_TSA = sum(BE_breakdown$twoPA + BE_breakdown$TPA)

# Three point shots attempted

BE_tpa_total = sum(BE_breakdown$TPA)

BE_tpa_total

# Reliance on three point shot 

BE_reliance = BE_tpa_total / BE_TSA

BE_reliance * 100

# Three point percentage

BE_TPper_total = median(BE_breakdown$TP_per)

BE_TPper_total * 100

# Offensive rebounds

BE_oreb_total = sum(BE_breakdown$oreb, na.rm = TRUE)

B10_oreb_total

# Free throws attempted

BE_FTA_total = sum(BE_breakdown$FTA)

BE_FTA_total

# Free throws made

BE_FTper_total = median(BE_breakdown$FT_per)

BE_FTper_total * 100

# Blocks per game

BE_BLKper_total = median(BE_breakdown$blk_per)

BE_BLKper_total 

# Steals per game

BE_STLper_total = median(BE_breakdown$stl_per)

BE_STLper_total 

```

```{r}
# PAC 12 Analysis

# Create desired data frame

PAC12_breakdown = filter(conference_breakdown, conf == "P12")

PAC12_breakdown


# Total Shots Attempted

PAC12_TSA = sum(PAC12_breakdown$twoPA + PAC12_breakdown$TPA)

# Three point shots attempted

PAC12_tpa_total = sum(PAC12_breakdown$TPA)

PAC12_tpa_total

# Reliance on three point shot 

PAC12_reliance = PAC12_tpa_total / PAC12_TSA

PAC12_reliance * 100

# Three point percentage

PAC12_TPper_total = median(PAC12_breakdown$TP_per)

PAC12_TPper_total * 100

# Offensive rebounds

PAC12_oreb_total = sum(PAC12_breakdown$oreb, na.rm = TRUE)

PAC12_oreb_total

# Free throws attempted

PAC12_FTA_total = sum(PAC12_breakdown$FTA)

PAC12_FTA_total

# Free throws made

PAC12_FTper_total = median(PAC12_breakdown$FT_per)

PAC12_FTper_total * 100

# Blocks per game

PAC12_BLKper_total = median(PAC12_breakdown$blk_per)

PAC12_BLKper_total 

# Steals per game

PAC12_STLper_total = median(PAC12_breakdown$stl_per)

PAC12_STLper_total 

```
```{r}
# SEC Analysis

# Create desired data frame

SEC_breakdown = filter(conference_breakdown, conf == "SEC")

SEC_breakdown


# Total Shots Attempted

SEC_TSA = sum(SEC_breakdown$twoPA + SEC_breakdown$TPA)

# Three point shots attempted

SEC_tpa_total = sum(SEC_breakdown$TPA)

SEC_tpa_total

# Reliance on three point shot 

SEC_reliance = SEC_tpa_total / SEC_TSA

SEC_reliance * 100

# Three point percentage

SEC_TPper_total = median(SEC_breakdown$TP_per)

SEC_TPper_total * 100

# Offensive rebounds

SEC_oreb_total = sum(SEC_breakdown$oreb, na.rm = TRUE)

SEC_oreb_total

# Free throws attempted

SEC_FTA_total = sum(SEC_breakdown$FTA)

SEC_FTA_total

# Free throws made

SEC_FTper_total = median(SEC_breakdown$FT_per)

SEC_FTper_total * 100

# Blocks per game

SEC_BLKper_total = median(SEC_breakdown$blk_per)

SEC_BLKper_total 

# Steals per game

SEC_STLper_total = median(SEC_breakdown$stl_per)

SEC_STLper_total 

```



# Summary


In my exploration of our NCAA college basketball data, I gained valuable insights into the intricacies of the game. Despite our collective experience in observing the dataset unfold during college basketball games, our analysis revealed an unexpected finding – the absence of a genuine correlation between counting statistics and draft picks. This contradicts the conventional belief that players excelling in points, rebounds, and assists have a higher chance of being drafted. Additionally, our examination indicated minimal variance between conferences concerning both traditional and advanced analytics. This lack of diversity can be attributed to the majority of games being played within a specific conference. Noteworthy observations emerged from our initial questions, particularly those addressing players with a low effective field goal percentage (eFG) and the conference with the highest eFG in 2018. These questions proved instrumental, leading us to explore correlations and significance within conferences and draft pick statuses.


My investigation into follow-up questions provided further depth to our understanding of college basketball dynamics. The diversity within college basketball, including varying roster sizes and play styles, adds complexity to the sport. Analyzing the number of players registered by each team over the past 13 years, I initially noted a left-skewed distribution, which upon closer inspection, was influenced by teams with incomplete data due to their recent entry into DI basketball. Removing these outliers revealed a more normal distribution. Surprisingly, schools with weaker basketball cultures registered more players, challenging the notion that prominent basketball schools require more replacements for departing players. Another intriguing aspect is the similarity in play styles among conferences with high draft pick production. Despite initial expectations of significant differences, the analysis revealed slim variations, suggesting a resemblance to the NBA's style of play. These findings enhance the complexity and uniqueness of college basketball, making it an interesting subject for further investigation.
